Test susieI using whole genome, s80.45 samples

### 2021-01-19 analysis/simulation-susieI-ukbWG-gtex.adipose_s80.45.Rmd
faster gwas

```{r parameters}
n.ori <- 40000 # number of samples
n <-  22542
p <- 656321 # number of SNPs
J <- 8021 # number of genes
```

The genotype data we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples. We then filtered samples based on relatedness, ethics and other qc metrics, that ended up with n = `r as.character(n)` samples.  We use SNP genotype data from chr 1 to chr 22 combined from UKB. SNPs are downsampled to 1/10 (randomly), eQTLs (see below for definition of eQTL) were added back. This ends up with p = `r as.character(p)` SNPs.

@@

Next, the analysis is done at the “region” level, which is 500kb bins along the genome. We also used LDetect to define regions. We are exploring several ways to select regions that contain true signals, e.g. based on regional sum of mr.ash PIP for genes/SNPs, region smallest TWAS p value for gene/SNPs, or regional bayes factors, etc.

4. Run susie iteratively
We then run susie for each of these regions. So the features of SuSiE are: SNPs and "genes" (not cis-eQTLs of that gene). We use the same prior for all SNPs and another prior for all "genes" when running SUSIE. In some settings, we also run SUSIE with null weight, which is calculated as `1- prior.SNP * n.SNP - prior.gene * n.gene`. We obtain the PIP for SNPs and gene in the region. After we run susie for all regions (one iteration), we take the average of all SNP PIPs as the prior of SNPs for the next iteration and similarly for the prior for genes.

5. We obtain PIP for genes from the last iteration as results.

```{r dirs}
simdatadir <- "~/causalTWAS/simulations/simulation_ashtest_20201001/"
outputdir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
susiedir <- "~/causalTWAS/simulations/simulation_susieI_20201001/"
simtag <- '20201001-1-3'
source('~/causalTWAS/causal-TWAS/code/gwas.R')
source('~/causalTWAS/causal-TWAS/code/ld.R')
source('~/causalTWAS/causal-TWAS/code/qqplot.R')
```

```{r causal}
exprgwasf <- paste0(simdatadir, simtag, ".exprgwas.txt.gz")
load(paste0(simdatadir, "simu_", simtag, "-pheno.Rd"))
caulist <- list()
for (chrom in 1:22) {
  load(paste0("~/causalTWAS/ukbiobank/ukb_chr", chrom ,"_s40.22.FBM.Rd"))
  load(paste0(simdatadir, "simu_s40000_GTEXadipose-B", chrom, "-cis-expr.Rd"))
  caulist[[chrom]]<- c(exprres$gnames[phenores$batch[[chrom]]$param$idx.cgene], dat$snp[phenores$batch[[chrom]]$param$idx.cSNP,])
}
cau <- unlist(caulist)
```
# Power estimation

We use gene.pve ~ 0.1, snp.pve ~ 0.5.

* For SNPs, we use $\pi_1 = 2.5e-3$, variance for effect size ~ $0.03^2$, power at 5e-8 p value cutoff:
```{r pow func} 
pow <- function(total, n, beta, cutp){
  rec <- rep(0, total)
  for (i in 1:total){
    x <- rnorm(n)
    y <- x * rnorm(1, sd = beta) + rnorm(n, sd = sqrt(2.5))
    lm.s <- lm(y~x)
    pv <- summary(lm.s)$coefficients[2,4]
    rec[i] <- pv
  }
  length(rec[rec < cutp])/length(rec)
}
```

```{r snp}
load("data/power_s40.22.Rd")
total <- 1e3
n <- 22542
#p1 <- pow(total, n, 0.0276, 5e-8)
print(p1)
```

* For genes, under low power setting, $\pi_1 = 0.05$, variance for effect size ~ $0.025^2$, power at 1e-5 cutoff:

```{r gene1}
#p2 <- pow(total, n, 0.025, 1e-5)
print(p2)
```

For genes, under high power setting, $\pi_1 = 0.02$, variance for effect size ~ $0.045^2$, power at 1e-5 cutoff:

```{r gene2}
#p3 <- pow(total, n, 0.045, 1e-5)
print(p3)
#save(p1,p2,p3, file = "data/power_s40.22.Rd")
```

# p value distribution

* TWAS p value of genes:

```{r gene twas, fig.height=8, fig.width=8}
chrom <- 1
a <- read.table(exprgwasf, header = T)
a$ifcausal <- ifelse(a$MARKER_ID %in% cau, 1, 0)
ax <- pretty(0:max(-log10(a$PVALUE)), n = 30)

par(mfrow=c(3,1))
h1 <- hist(-log10(a$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey", xlim= c(3,20), ylim =c(0,50)); grid()
h2 <- hist(-log10(a[a$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon", xlim= c(3,20), ylim =c(0,50));grid()

cat("number of genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5, ]))
cat("number of causal genes p < 1e-5:", nrow(a[a$PVALUE < 1e-5 & a$ifcausal ==1, ]))

plot(a[a$X.CHROM ==chrom, ]$BEGIN, -log10(a[a$X.CHROM ==chrom, ]$PVALUE), col = a[a$X.CHROM ==chrom, ]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
points(a[a$X.CHROM ==chrom & a$ifcausal ==1, ]$BEGIN, -log10(a[a$X.CHROM ==chrom & a$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
grid()
```

* qq plot for genes
```{r qqgene, fig.height= 5, fig.width=5}
gg_qqplot(a$PVALUE) +
  theme_bw(base_size = 24) +
  theme(
    axis.ticks = element_line(size = 0.5),
    panel.grid = element_blank()
    # panel.grid = element_line(size = 0.5, color = "grey80")
  )
```

*  p value of SNPs (GWAS):

```{r snp twas, fig.height=8, fig.width=8}
snpgwasf <- paste0(simdatadir, simtag, ".snpgwas.txt.gz")

b <- fread(snpgwasf, header = T)
b$ifcausal <- ifelse(b$MARKER_ID %in% cau, 1, 0)
ax <- pretty(0:max(-log10(b$PVALUE)), n = 30)

par(mfrow=c(3,1))
h1 <- hist(-log10(b$PVALUE), breaks = 100, xlab = "-log10(p)", main = "P value distribution-all", col = "grey", xlim= c(3,20), ylim =c(0,100)); grid()
h2 <- hist(-log10(b[b$ifcausal == 1, ]$PVALUE), breaks = h1$breaks, xlab = "-log10(p)", main = "P value distribution-causal", col = "salmon", xlim= c(3,20), ylim =c(0,100));grid()

cat("number of SNPs < 5e-8: ", nrow(b[b$PVALUE < 5e-8,]))

plot(b[b$X.CHROM ==chrom, ]$BEGIN, -log10(b[b$X.CHROM ==chrom, ]$PVALUE), col = b[b$X.CHROM ==chrom,]$ifcausal + 1, xlab = paste0("chr", chrom), ylab = "-log10(pvalue)")
points(b[b$X.CHROM ==chrom & b$ifcausal ==1, ]$BEGIN, -log10(b[b$X.CHROM ==chrom & b$ifcausal ==1, ]$PVALUE), col = "red", pch =19)
grid()
```

* qq plot for SNPs
```{r qqsnp, fig.height= 5, fig.width=5}
gg_qqplot(b$PVALUE) +
  theme_bw(base_size = 24) +
  theme(
    axis.ticks = element_line(size = 0.5),
    panel.grid = element_blank()
    # panel.grid = element_line(size = 0.5, color = "grey80")
  )
```
# Block features

```{r block}

a <- read.table(paste0(outputdir, "20201001-1-1.config9.gene.nofilter.r.txt"), header =T)
b <-  read.table(paste0(outputdir, "20201001-2-1.config9.gene.nofilter.r.txt"), header =T)
m1 <- read.table(paste0(outputdir, "20201001-1-1.config9.nofilter.r.txt"), header =T)
m2 <- read.table(paste0(outputdir, "20201001-2-1.config9.nofilter.r.txt"), header =T)

hist(a$p1-a$p0, main = NULL, xlab = "Block size(bp)", col = "salmon")

par(mfrow=c(2,2))
hist(a$nCausal, xlab = "No.causal genes", col = "salmon", breaks=100, xlim=c(0,15), main ="low power")
hist(b$nCausal, xlab = "No.causal genes", col = "salmon", breaks=100, xlim=c(0,15), main ="high power")
hist(m1$nCausal, xlab = "No.causal genes + SNPs", col = "salmon", breaks=100, xlim=c(0,15), main ="low power")
hist(m2$nCausal, xlab = "No.causal genes + SNPs", col = "salmon", breaks=100, xlim=c(0,15), main ="high power")
```

# TWAS FDP

```{r PIPplot functions}

library(ggplot2)
library(cowplot)
library(plotrix)

.obn <- function(pips, ifcausal, mode = c("PIP", "FDR")){
  a_bin <- cut(pips, breaks= seq(0, 1, by=0.1))
  if (mode == "PIP") {
    ob = c(by(ifcausal, a_bin, FUN = sum))
  } else if (mode == "FDR"){
    ob = c(by((1-ifcausal), a_bin, FUN = sum))
  }
  return(ob)
}

nca_plot <- function(pips, ifcausal, runtag = NULL, mode = c("PIP", "FDR"), main = mode[1], ...){

  # ifcausal:0,1, runtag: for adding std.
  if (is.null(runtag)){
     se = 0 
  } else{
    dflist <- list()
    for (rt in unique(runtag)){
      pips.rt <- pips[runtag == rt]
      ifcausal.rt <- ifcausal[runtag == rt]
      dflist[[rt]] <- .obn(pips.rt, ifcausal.rt, mode = mode)
    }
    nca_mean <- colMeans(do.call(rbind, dflist))
    se <- apply(do.call(rbind, dflist), 2, plotrix::std.error)
  }

  df <- data.frame("ncausal" = nca_mean, "se" = se)

  fig <- ggplot(df) +
    geom_bar( aes(x=seq(0, 1, by=0.1)[1:10]+0.05, y=ncausal), color ="black", stat="identity", fill="salmon", alpha=0.7, width = 0.1) +
    geom_errorbar( aes(x= seq(0, 1, by=0.1)[1:10] + 0.05, ymin= ncausal-se, ymax=ncausal+se), width=0.05, colour="black", alpha=0.9, size=0.5) +
    ggtitle(main) +
    xlab("PIP") + ylab("No. causal genes")
    theme_cowplot()
  return(fig)
}

ncausal_plot <- function(pipfs, format = "susie", main = "PIP"){
  df <- NULL
  for (i in 1:length(pipfs)) {
    res <- fread(pipfs[i], header = T)
    res <- data.frame(res[res$type  =="gene", ])
    res$runtag <- i
    if (format == "susie"){
       res <- rename(res, c("susie_pip" = "pip") )
    } else if (format == "SER"){
       res <- rename(res, c("SERpip" = "pip"), )
    }
    res <- res[complete.cases(res),]
    df <- rbind(df, res)
  }

  fig <- nca_plot(df$pip, df$ifcausal, df$runtag, mode ="PIP", main = main)
  return(fig)
}

.exob <- function(pips, ifcausal, mode = c("PIP", "FDR")){
  a_bin <- cut(pips, breaks= seq(0, 1, by=0.1))
  if (mode == "PIP") {
    ex = c(by(pips, a_bin, FUN = mean))
    ob = c(by(ifcausal, a_bin, FUN = mean))
  } else if (mode == "FDR"){
    ex = c(by(pips, a_bin, FUN = mean))
    ob = 1 - c(by(ifcausal, a_bin, FUN = mean))
  }
  return(list("expected" = ex, "observed" = ob))
}

dot_plot = function(df, main) {
        ggplot(df, aes(x=mean_pip, y=observed_freq)) + 
          geom_errorbar(aes(ymin=observed_freq-se, ymax=observed_freq+se), colour="black", size = 0.5, width=.01) +
          geom_point(size=1.5, shape=21, fill="#002b36") + # 21 is filled circle
          xlab("Expected") +
          ylab("Observed") +
          coord_cartesian(ylim=c(0,1), xlim=c(0,1)) +
          geom_abline(slope=1,intercept=0,colour='red', size=0.2) +
          ggtitle(main) +
          expand_limits(y=0) +                        # Expand y range
          theme_cowplot()}

cp_plot <- function(pips, ifcausal, runtag = NULL, mode = c("PIP", "FDR"), main = mode[1], ...){

  # ifcausal:0,1, runtag: for adding std.
  if (is.null(runtag)){
     se = 0 
  } else{
    dflist <- list()
    for (rt in unique(runtag)){
      pips.rt <- pips[runtag == rt]
      ifcausal.rt <- ifcausal[runtag == rt]
      dflist[[rt]] <- .exob(pips.rt, ifcausal.rt, mode = mode)
    }
    mean_pip <- colMeans(do.call(rbind, lapply(dflist, '[[', "expected")))
    observed_freq <- colMeans(do.call(rbind, lapply(dflist, '[[', "observed")))
    se <- apply(do.call(rbind, lapply(dflist, '[[', "observed")), 2, plotrix::std.error)
  }

  df <- data.frame("mean_pip" = mean_pip, "observed_freq"= observed_freq, "se" = se)
  dot_plot(df, main = main)

  #plot(Expected, Observed, xlim= c(0,1), ylim=c(0,1), pch =19, main = main, ...)
  #lines(x = c(0,1), y = c(0,1), col ="grey", lty = 2)
}

caliPIP_plot <- function(pipfs, format = "susie", main = "PIP"){
  df <- NULL
  for (i in 1:length(pipfs)) {
    res <- fread(pipfs[i], header = T)
    res <- data.frame(res[res$type  =="gene", ])
    res$runtag <- i
    if (format == "susie"){
       res <- rename(res, c("susie_pip" = "pip"))
    } else if (format == "SER"){
       res <- rename(res, c("SERpip" = "pip"))
    }
    res <- res[complete.cases(res),]
    df <- rbind(df, res)
  }

  fig <- cp_plot(df$pip, df$ifcausal, df$runtag, mode ="PIP", main = main)
  return(fig)
}

# pipfs is just for ifcausal info
caliFDP_plot <- function(gwasfs, pipfs,  format= "susie",  main = "FDP"){
  df <- NULL
  for (i in 1:length(pipfs)) {
    pipres <- fread(pipfs[i], header = T)
    pipres <- data.frame(pipres[pipres$type  =="gene", ])
    pipres$runtag <- i
    if (format == "susie"){
       pipres <- rename(pipres, c("susie_pip" = "pip"))
    } else if (format == "SER"){
       pipres <- rename(pipres, c("SERpip" = "pip"))
    }
    gwasres <- read.table(gwasfs[i], header = T)
    gwasres <- rename(gwasres, c("MARKER_ID" = "name"))
    gwasres$FDR <- p.adjust(gwasres$PVALUE, method = "fdr")
    res <- merge(gwasres, pipres, by = "name", all = T)
    res <- res[complete.cases(res),]
    df <- rbind(df, res)
  }
  fig <- cp_plot(df$FDR, df$ifcausal, df$runtag, mode ="FDR", main = main)
  cat("FDP at bonferroni corrected p = 0.05: ", 1 - mean(df[df$PVALUE < 0.05 /dim(df)[1], "ifcausal"]))
  return(fig)
}

scatter_plot_PIP_p <- function(pipfs, gwasfs, pipformat = "susie", main ="PIP-p"){
  df <- NULL
  for (i in 1:length(pipfs)) {
    pipres <- fread(pipfs[i], header = T)
    pipres <- data.frame(pipres[pipres$type  =="gene", ])
    pipres$runtag <- i
    if (pipformat == "susie"){
       pipres <- rename(pipres, c("susie_pip" = "pip"))
    } else if (pipformat == "SER"){
       pipres <- rename(pipres, c("SERpip" = "pip"))
    }
    gwasres <- read.table(gwasfs[i], header = T)
    gwasres <- rename(gwasres, c("MARKER_ID" = "name"))
    res <- merge(gwasres, pipres, by = "name", all = T)
    res <- res[complete.cases(res),]
    df <- rbind(df, res)
  }

  df <- rename(df, c( "PVALUE" = "TWAS.p"))
  df[,"TWAS.p"] <- -log10(df[, "TWAS.p"])

  df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1),to=c("darkgreen", "salmon"))
  plot(df$TWAS.p, df$pip, col = df$ifcausal, main = main, xlab = "-log10(TWAS p value)", ylab = "PIP")

  # df$ifcausal <- mapvalues(df$ifcausal, from=c(0,1), to=c("Non causal", "Causal"))
  # fig <- plot_ly(data = df, x = ~ TWAS.p, y = ~ pip, color = ~ ifcausal,
  #                colors = c( "salmon", "darkgreen"), type ="scatter", text = ~ paste("Name: ", paste0(runtag,":",name),
  #                                                                  "\nChr: ", chr,  "\nPos:", pos))

}
```

```{r TWAS FDP, fig.height=4, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
gwasfs <-  paste0(simdatadir, "20201001-", tags, ".exprgwas.txt.gz")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")

f1 <- caliFDP_plot(gwasfs[1:5], susieIfs2[1:5], format= "susie", main = "TWAS FDP (low power)")
f2 <- caliFDP_plot(gwasfs[6:10], susieIfs2[6:10], format= "susie", main ="TWAS FDP (high power)")

gridExtra::grid.arrange(f1, f2, ncol =2)
```

# L's effect on SuSiE

We run SuSiE on each block with true priors and true prior variances. We tested the difference between using `L=1` and `L=5`. We only did this for the low power setting.

```{r reg functions}
ncs_ncausal <- function(susief, main = "nCSvs.nCausal") {
  dt <- fread(susief, header = T)
  ncausal <- dt[, sum(ifcausal), by=list(b,rn)]
  nCS <- dt[, max(cs_index), by=list(b,rn)]
  plot(jitter(ncausal$V1), jitter(nCS$V1), xlab= "No.causal/block", ylab = "No. credible set/block", main= main)
  return(dt)
}

sumPIP_ncausal <- function(susief, main = "sumPIPvs.nCausal"){
  dt <- fread(susief, header = T)
  ncausal <- dt[, sum(ifcausal), by=list(b,rn)]
  sumpip <- dt[, sum(susie_pip), by=list(b,rn)]
  plot(jitter(ncausal$V1), sumpip$V1, xlab= "No.causal/block", ylab = "Regional sum PIP", main = main)
  return(dt)
}
```
## L=1

Note: the data has been added some noise (both X and Y axes for credible set vs. no.causal, X axis for sum of PIP vs. no.causal) for visualization purposes. 
```{r L=1, fig.width=8, fig.height=4}
tags <- paste(rep(1, each = 5), 1:5, sep = "-")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config11.susieI.txt")
f1 <- caliPIP_plot(susieIfs2, main = "L=1")
f2 <- ncausal_plot(susieIfs2, format = "susie", main = "No. Causal Genes")
gridExtra::grid.arrange(f1, f2, ncol =2)

par(mfrow=c(1,2))
dt <- sumPIP_ncausal(susieIfs2[2], main = "L=1")
dt <-ncs_ncausal(susieIfs2[2], main = "L=1, CS=0.95")
```

* True parameters: gene $\pi_1$, `r g <- nrow(dt[type =="gene" & ifcausal !=0])/ nrow(dt[type =="gene"]);g`, SNP $\pi_1$, `r s <-nrow(dt[type =="SNP" & ifcausal !=0])/ nrow(dt[type =="SNP"]);s`. [enrichment = `r g/s`].
* Average PIP: gene `r g <- mean(dt[type =="gene", susie_pip]);g`, SNP `r s <-mean(dt[type =="SNP", susie_pip]);s` . [enrichment = `r g/s`].
* Proportion in credible set (0.95): gene `r g <-nrow(dt[type =="gene" & cs_index !=0])/ nrow(dt[type =="gene"]);g `, SNP `r s <-nrow(dt[type =="SNP" & cs_index !=0])/ nrow(dt[type =="SNP"]);s`. [enrichment = `r g/s`].

## L=5

Note: the data has been added some noise (both X and Y axes for credible set vs. no.causal, X axis for sum of PIP vs. no.causal) for visualization purposes.

```{r L=5, fig.width=8, fig.height=4}
tags <- paste(rep(1, each = 5), 1:5, sep = "-")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config12.susieI.txt")

f1 <- caliPIP_plot(susieIfs2, main = "L=5")
f2 <- ncausal_plot(susieIfs2, format = "susie", main = "No. Causal Genes")
gridExtra::grid.arrange(f1, f2, ncol =2)

par(mfrow=c(1,3))
dt <- sumPIP_ncausal(susieIfs2[2], main = "L=5")
dt <- ncs_ncausal(susieIfs2[2], main = "L=5, CS=0.95")

susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config13.susieI.txt")
dt2 <- ncs_ncausal(susieIfs2[2], main = "L=5, CS=0.8")
```

* True parameters: gene $\pi_1$, `r g <- nrow(dt[type =="gene" & ifcausal !=0])/ nrow(dt[type =="gene"]);g`, SNP $\pi_1$, `r s<-nrow(dt[type =="SNP" & ifcausal !=0])/ nrow(dt[type =="SNP"]);s`. [enrichment = `r g/s`].
* Average PIP: gene `r g <- mean(dt[type =="gene", susie_pip]);g`, SNP `r s <-mean(dt[type =="SNP", susie_pip]);s` . [enrichment = `r g/s`].
* Proportion in credible set (0.95): gene `r g <-nrow(dt[type =="gene" & cs_index !=0])/ nrow(dt[type =="gene"]);g `, SNP `r s <-nrow(dt[type =="SNP" & cs_index !=0])/ nrow(dt[type =="SNP"]);s`. [enrichment = `r g/s`].
* Proportion in credible set (0.8): gene `r g <-nrow(dt2[type =="gene" & cs_index !=0])/ nrow(dt2[type =="gene"]);g `, SNP `r s <-nrow(dt2[type =="SNP" & cs_index !=0])/ nrow(dt2[type =="SNP"]);s`. [enrichment = `r g/s`].

@@

```{r param functions}
show_param <- function(phenofs, susieIfs, susieIfs2){
  pars <- do.call(rbind, lapply(phenofs, function(x) {load(x); 
    c(phenores$param$pve.gene.truth,
      phenores$param$pve.snp.truth,
      length(phenores$batch[[1]]$param$idx.cgene)/phenores$batch[[1]]$param$J,
      length(phenores$batch[[1]]$param$idx.cSNP)/phenores$batch[[1]]$param$M)}))

@@

  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(prior.gene_rec[prior.gene_rec!=0], 1), tail(prior.SNP_rec[prior.SNP_rec!=0],1))}))

@@

  st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"]/df[,"pi1.SNP_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]/df[1,"pi1.SNP_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, min(max(dfp[,1],t) *1.05, 150)))
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=2:3, 0, labels = c("susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h= t, lty = 2, col= "darkgreen", lwd=1.5)
  grid()
}

gpip_dist <- function(susiefs, ...){
  dflist <- list()
  for (f in susiefs){
    dflist[[f]] <- read.table(f, header =T , stringsAsFactors = F)
  }
  df <- do.call(rbind, dflist)
  hist(df[df$type == "gene", "susie_pip"], xlab = "gene susie PIP", 
       breaks = 50, ylim = c(0,20), xlim=c(0,1), col = "salmon", ...)
}
```

## susieI (1)
* Regions: all regions, 500kb uniform regions.
* Susie run parameters: `L=1`.  We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). Prior variance and residual variance were calculated by SUSIE for each region.

```{r param1susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
gwasfs <-  paste0(simdatadir, "20201001-", tags, ".exprgwas.txt.gz")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config1.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config1.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

```{r pip1susie, fig.height=6, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
```

## susieI (2)
* Regions: all regions, LD-defined regions.
* Susie run parameters: `L=1`.  We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). Prior variance and residual variance were calculated by SUSIE for each region.

```{r param6susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config6.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config6.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

```{r pip6susie, fig.height=6, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
```

## susieI (3)
* Regions: all regions, LD-defined regions.
* Susie run parameters: `L=10`.  We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes). Prior variance and residual variance were calculated by SUSIE for each region.

```{r param8susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config8.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config8.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

```{r pip8susie, fig.height=6, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
```

## susieI (4)
* Regions: all regions, LD-defined regions.
* Susie run parameters: `L=1`. We use true prior variance and plug into SUSIE, so this should be the same as SER/EM (2). We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).

```{r param9susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config9.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config9.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

```{r pip9susie, fig.height=6, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high power)")
gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
```

## susieI (5)
* Regions: all regions, LD-defined regions.
* Susie run parameters: `L=5`. We use true prior variance and plug into SUSIE. We initialize with prior for genes and SNPs as uniform. gene.pve ~ 0.1, snp.pve ~ 0.5.  Null weight is calculated based on prior of genes and SNPs ( 1 - sum of priors for snps and genes).

```{r param10susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(simdatadir, "simu_20201001-", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, "20201001-", tags, ".config10.susieIres.Rd")
susieIfs2 <- paste0(outputdir, "20201001-", tags, ".config10.susieI.txt")

df <- show_param(phenofs, susieIfs, susieIfs2)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```

```{r pip10susie, fig.height=6, fig.width=8}
f1 <- caliPIP_plot(susieIfs2[1:5], format ="susie", main = "PIP(low power)" )
f2 <- caliPIP_plot(susieIfs2[6:10], format ="susie", main = "PIP(high power)" )
f3 <- ncausal_plot(susieIfs2[1:5], format = "susie", main = "No. Causal Genes (low power)")
f4 <- ncausal_plot(susieIfs2[6:10], format = "susie", main = "No. Causal Genes (high poweunr)")
gridExtra::grid.arrange(f1, f2, f3, f4, ncol =2)
```

>>>>

```{r parameters}
n.ori <- 80000 # number of samples
n <-  45542
p <- 6656321 # number of SNPs
J <- 8021 # number of genes
```

The genotype data we used is from UKB biobank, randomly selecting `r as.character(n.ori)` samples. We then filtered samples based on relatedness, ethics and other qc metrics, that ended up with n = `r as.character(n)` samples.  We use SNP genotype data from chr 1 to chr 22 combined from UKB. There total = `r as.character(p)` SNPs.

@@

Next, the analysis is done at the “region” level. We used LDetect to define regions.

4. Run ctwas_rss
We downsampled to 1/10 of SNPs to run ctwas_rss

```{r dirs}
library(ctwas)
library(data.table)
source('~/causalTWAS/causal-TWAS/code/qqplot.R')
outputdir <- "~/causalTWAS/simulations/simulation_ctwas_rss_20210117/"
runtag <- 'ukb-s80.45-adi'
```

@@

```{r param functions}
show_param <- function(phenofs, susieIfs, susieIfs2, thin = 1){
  pars <- do.call(rbind, lapply(phenofs, function(x) {load(x); 
    c(phenores$param$pve.gene.truth,
      phenores$param$pve.snp.truth,
      length(phenores$batch[[1]]$idx.cgene)/phenores$batch[[1]]$J,
      length(phenores$batch[[1]]$idx.cSNP)/phenores$batch[[1]]$M)}))

@@

  param.s <- do.call(rbind, lapply(susieIfs, function(x) {load(x); c(tail(group_prior_rec[1,][!is.na(group_prior_rec[1,])], 1), tail(group_prior_rec[2,][!is.na(group_prior_rec[1,])], 1) * thin)}) )

@@

  df[,"pi1.SNP_susietruth"] <- 1
  st <-  cbind(df[,"pi1.gene_susietruth"]/df[,"pi1.SNP_susietruth"], 1:nrow(df), 2 + 1:nrow(df)/nrow(df)/3)
  s <- cbind(df[,"pi1.gene_susieI"]/df[,"pi1.SNP_susieI"], 1:nrow(df), 3 + 1:nrow(df)/nrow(df)/3)
  t <- df[1,"pi1.gene_truth"]/df[1,"pi1.SNP_truth"]
  dfp <- rbind(st,s)
  plot(dfp[,3], dfp[,1], col = dfp[,2], pch = 19, ylab = "Enrichment (gene/snp)", xaxt = "n", xlab="", xlim = c(0.8, 3.5),frame.plot=FALSE, ylim = c(0, min(max(dfp[,1],t) *1.05, 150)))
  axis(side=1, at=1:2, labels = FALSE, tick = F)
  text(x=2:3, 0, labels = c("susieI_truth", "susieI"), xpd = T, pos =1)
  abline(h= t, lty = 2, col= "darkgreen", lwd=1.5)
  grid()
}

```

## ctwas_rss (1)
* run parameters: downsample 1/10.

```{r param1susie, fig.height=6, fig.width=8}
tags <- paste(rep(1:2, each = 5), 1:5, sep = "-")
phenofs <- paste0(outputdir, runtag, "_simu", tags, "-pheno.Rd")
susieIfs <- paste0(outputdir, runtag, "_simu", tags, "_config1.s2.susieIrssres.Rd")
susieIfs2 <- paste0(outputdir, runtag, "_simu",tags, "_config1.s2.susieIrss.txt")

df <- show_param(phenofs, susieIfs, susieIfs2, thin = 0.1)
par(mfrow = c(2,3))
plot_param(df[1:5,], main = "low power")
plot_param(df[6:10,], main = "high power")
```
